---
title: "notebook_behav"
author: "Marie"
date: "`r Sys.Date()`"
output: html_document
---

# ----- R UNIVERSE PREP -----

```{r setup}
rm(list = ls()) # remove everything old that might be in my var space
knitr::opts_chunk$set(echo = TRUE) # this is to connect markdown and r - i don't think i need this but whatever
pacman::p_load(tidyverse, nlme, CorrMixed, lme4, lmerTest, geepack, ggplot2, geomtextpath, car, emmeans, effectsize, simr, matrixStats) # pacman is a package manager for R (!!! useful). Install packages if missing and load them 
options(scipen=999) # strongly avoid scientific notations (none of the e bs)
```

# ----- LOAD DATA -----

```{r data}
data_eo <- read.table("/Users/abhishekanand/Documents/OTPH_TF/dat_EO_relPower_withbehav_missval.csv", header=T, sep=",")
data_ec <- read.table("/Users/abhishekanand/Documents/OTPH_TF/dat_EC_relPower_withbehav_missval.csv", header=T, sep=",")

data_eo$Drug <- factor(data_eo$Drug)
data_eo$TP <- factor(data_eo$TP)
data_eo$id <- factor(data_eo$id)

data_ec$Drug <- factor(data_ec$Drug)
data_ec$TP <- factor(data_ec$TP)
data_ec$id <- factor(data_ec$id)

# SPLIT DATA between ot and pl

data_eo_ot_all<-data_eo[data_eo$Drug==1,]
data_eo_pl_all<-data_eo[data_eo$Drug==2,]

data_ec_ot_all<-data_ec[data_ec$Drug==1,]
data_ec_pl_all<-data_ec[data_ec$Drug==2,]
```

# ----- Min worlking example: theta, alertness -----

```{r minexample-theta-alertness-MODEL}

# LMM
# --> step 1
model = lmer(Theta ~ Theta_base + TP*Drug*Alertness_reversed + (1|id), data=data_ec)
#tab_model(model , digits = 4)
print(summary(model ))

# anova
print(anova(model, type=3))

# --> step 2: only look at ot separately and pl separately
# model = lmer(Theta ~ Theta_base + TP*Alertness_reversed + (1|id), data=data_ec_otonly)
# look at trends

# --> step 3: only look at each timepoint separately 
# model = lmer(Theta ~ Theta_base + Alertness_reversed + (1|id), data=data_ec_otonly_tp1)


```

# ============================================================================
# ===============================ANALYZE======================================
# ============================================================================

# ----> 1. Run LMMs iteratively, look for main/interaction effects, that defines what to look at in more detail <----

```{r omnibus, echo=FALSE}

omnibus_found <- list()

path_to_savefile = "//Users/abhishekanand/Documents/OTPH_TF/RR/RR_OTPH_results1_omnibus.txt"
sink(path_to_savefile)

# LOOP THROUGH EYE OPEN/CLOSED
for (eyes_cond in c("eyes_open", "eyes_closed")){
  
  # chose the right data depending on eyes cond
  if (eyes_cond == "eyes_open") {
    mydata_main   <- data_eo
    mydata_ot_all <- data_eo_ot_all
    mydata_pl_all <- data_eo_pl_all
  } else {
    mydata_main   <- data_ec
    mydata_ot_all <- data_ec_ot_all
    mydata_pl_all <- data_ec_pl_all
  } 

  # LOOP THROUGH BANDS
  bands <- c("Theta", "Alpha", "Beta")
  for (band in bands) {
  
    # LOOP THROUGH BEHAV MEASURES
    behav_vars <- c("Alertness_reversed", "Excitement_reversed", "Sociability_reversed")
    for (behav_var in behav_vars) {
      
      # do the analysis for this eyes condition, band, measure
      cat("\n-------------------------------------------------------------------------------------------------------------\n")
      cat("\n-------------------------------------------EYES = ", eyes_cond, "--------------------------------------------\n")
      cat("\n----------------------------------------------BAND = ", band, "----------------------------------------------\n")
      cat("\n------------------------------------------MEASURE = ", behav_var, "------------------------------------------\n")
      cat("\n-------------------------------------------------------------------------------------------------------------\n")
      
      cat("\n===== Anova results, after fitting lmm =====\n\n")
      
      formula_str <- paste0(band, " ~ ", band, "_base + TP*Drug*", behav_var, " + (1|id)")
      # save formula in output txt
      cat("\n\n")
      cat(formula_str)
      cat("\n\n")
      model <- lmer(as.formula(formula_str), data = mydata_main)  
      a <- anova(model, type=3)
      print(a)
      
      # if anova shows significant main or interaction --> update list of significant results
      
      effects <- c(behav_var, paste0("TP:", behav_var), paste0("Drug:", behav_var), paste0("TP:Drug:", behav_var))      
      for (effect in effects) {
        if (a[effect, "Pr(>F)"] < 0.05) {
          omnibus_found <- c(
            omnibus_found,
            sprintf(
              "%s: %s, %s --> signif: %s -- %s",
              eyes_cond, band, behav_var, effect, a[effect, "Pr(>F)"]))}}
  }}}
cat("\n=========================================================\n\n")
cat("===== Found significat results, to explore further: =====\n\n")
cat("=========================================================\n\n")
print(omnibus_found)
cat("=========================================================\n\n")

#return control back to console
sink()
```

# ----> Whever we found an effect in the main LMM, let's explore that further <----
# run separate LMMs for oxytocin and placebo, estimate slopes (then split into time points)

# 2. POSTHOCS - LMMs for PL/OT separately

```{r lmm-ot-pl-EO, echo=FALSE}

# function to do a model and check the slope
check_slopes <- function(dat, band, behav_var) {
  # lmm
  formula_str <- paste0(band, " ~ ", band, "_base + TP*", behav_var, " + (1|id)")
  # print formula
  cat("\n\n")
  cat(formula_str)
  cat("\n\n")
  model <- lmer(as.formula(formula_str), data = dat)  
  # anova
  myanova <- anova(model, type=3)
  # estimate slopes
  trend <- emtrends(model,  as.formula("~1"),var = behav_var)
  summary_trend_df<- as.data.frame(test(trend))
  summary_trend_df <- summary_trend_df %>%
        mutate(t = .data[[paste0(behav_var, ".trend")]] / SE)
  r_trend <- t_to_r(summary_trend_df$t, summary_trend_df$df)
  summary_trend_df$r <- r_trend$r
  myslopes = as.matrix(summary_trend_df)
  return(list(anova_return = myanova, slopes = myslopes))}

# -----------------------------------------------------------------------------

# ----------- EO data ----------- 

path_to_savefile = "//Users/abhishekanand/Documents/OTPH_TF/RR/RR_OTPH_results2_posthoc_ot_pl_slopes_EO.txt"
sink(path_to_savefile)

cat("\n\n                  EYES OPEN")

mydata_ot <- data_eo_ot_all
mydata_pl <- data_eo_pl_all

# combinations where we found sign in EYES OPEN

effect_pairs_eo <- list(
  c("Theta", "Alertness_reversed"),
  c("Beta", "Alertness_reversed"),
  c("Theta", "Excitement_reversed"),
  c("Alpha", "Excitement_reversed"),  
  c("Beta", "Excitement_reversed"),
  c("Alpha", "Sociability_reversed"),
  c("Beta", "Sociability_reversed"))

# iterate through EO effects
for (e in effect_pairs_eo) {

  bband <- e[1]
  bbehav <- e[2]
  
  cat("\n\n=========================================================\n\n")
  cat(paste0("      MENTAL STATE: ", bbehav, "; BAND: ", bband))
  cat("\n\n=========================================================\n\n")
  
  # OT and PL separately
  out_ot <- check_slopes(dat=mydata_ot, band=bband, behav_var=bbehav)
  out_pl <- check_slopes(dat=mydata_pl, band=bband, behav_var=bbehav)
  
  cat("\n\nANOVA Oxytocin:\n\n")
  print(out_ot$anova)
  cat("\n\nANOVA Placebo:\n\n")
  print(out_pl$anova)    
  
  cat("\n\nSlopes Oxytocin:\n\n")
  print(out_ot$slopes)
  cat("\n\nSlopes Placebo:\n\n")
  print(out_pl$slopes)  
  
}

sink()

# ----------- EC data ----------- 

path_to_savefile = "//Users/abhishekanand/Documents/OTPH_TF/RR/RR_OTPH_results2_posthoc_ot_pl_slopes_EC.txt"
sink(path_to_savefile)

cat("\n\n                  EYES CLOSED")

mydata_ot <- data_ec_ot_all
mydata_pl <- data_ec_pl_all

# combinations where we found sign in EYES OPEN

effect_pairs_ec <- list(
  c("Theta", "Alertness_reversed"),
  c("Alpha", "Sociability_reversed"),
  c("Beta", "Sociability_reversed"))

# iterate through EO effects
for (e in effect_pairs_ec) {

  bband <- e[1]
  bbehav <- e[2]
  
  cat("\n\n=========================================================\n\n")
  cat(paste0("      MENTAL STATE: ", bbehav, "; BAND: ", bband))
  cat("\n\n=========================================================\n\n")
  
  # OT and PL separately
  out_ot <- check_slopes(dat=mydata_ot, band=bband, behav_var=bbehav)
  out_pl <- check_slopes(dat=mydata_pl, band=bband, behav_var=bbehav)
  
  cat("\n\nANOVA Oxytocin:\n\n")
  print(out_ot$anova)
  cat("\n\nANOVA Placebo:\n\n")
  print(out_pl$anova)    
  
  cat("\n\nSlopes Oxytocin:\n\n")
  print(out_ot$slopes)
  cat("\n\nSlopes Placebo:\n\n")
  print(out_pl$slopes)  
  
}

sink()
```

# 3. POSTHOCS - slopes by timepoint
# At which timepoints is the Alertness slope significant?

```{r tp-EO, echo=True}

# ============================================================================
# =============================POSTHOC - TPs======================================
# ==============================EYES-OPEN=====================================
# ============================================================================

path_to_savefile <- "//Users/abhishekanand/Documents/OTPH_TF/RR/RR_OTPH_results3_posthoc_bytimepoint_EOonly.txt"

con <- file(path_to_savefile, "w")

cat("\n\n                  EYES OPEN\n", file = con)

for (e in effect_pairs_eo) {

  bband <- e$band
  bbehav <- e$behav
  ddat <- e$dat

  formula_str <- paste0(
    bband, " ~ ", bband, "_base + TP*", bbehav, " + (1|id)"
  )

  model <- lmer(as.formula(formula_str), data = ddat)

  trends_tp <- emtrends(
    model,
    as.formula("~TP"),
    var = bbehav
  )

  tp_df <- as.data.frame(test(trends_tp))

  tp_df <- tp_df %>%
    mutate(
      t = .data[[paste0(bbehav, ".trend")]] / SE
    )

  r_trend <- t_to_r(tp_df$t, tp_df$df)
  tp_df$r <- r_trend$r

  tp_out <- tp_df %>%
    select(
      TP,
      all_of(paste0(bbehav, ".trend")),
      SE,
      df,
      t,
      p.value,
      r
    )

  # Formula
  cat("\n\n", formula_str, "\n\n", file = con)

  # Column headers
  cat(
    "TP    ", bbehav, ".trend    SE       df       t       p.value      r\n",
    file = con
  )

  # Values
  cat(
    paste(
      sprintf(
        "%-5s %-17.3f %-8.3f %-8.2f %-7.2f %-10.3f %.3f",
        tp_out$TP,
        tp_out[[paste0(bbehav, ".trend")]],
        tp_out$SE,
        tp_out$df,
        tp_out$t,
        tp_out$p.value,
        tp_out$r
      ),
      collapse = "\n"
    ),
    "\n",
    file = con
  )
}

close(con)
```
